---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Needed data is loaded by

```{r load_data,echo=TRUE}
library(data.table)
library(knitr)

act<-read.csv("activity.csv",header = TRUE,stringsAsFactors = FALSE,na.strings = "NA")
act$date<-as.Date(act$date,"%Y-%m-%d")
act<-data.table(act)
```

## What is mean total number of steps taken per day?
Means and medians are presented in the next table with the respective histogram

```{r means_per_day,echo=TRUE}
pasosdia<-act[,sum(steps),by=date]
hist(pasosdia$V1,breaks=20,col=terrain.colors(length(unique(pasosdia$date))))
act[,mean(steps,na.rm = TRUE),by=date]
act[,as.double(median(steps,na.rm = TRUE)),by=date]
```

## What is the average daily activity pattern?
The time series plot is presented in next figure

```{r daily_activity,echo=TRUE}
pasosinterv<-act[,mean(steps,na.rm = TRUE),by=interval]
plot(pasosinterv$interval,pasosinterv$V1,type="l")
```

Max value is 
```{r max_value,echo=TRUE}
mayor<-which.max(pasosinterv$V1)
pasosinterv[mayor,]
```


## Imputing missing values

The total of missing values in the base is given by
```{r NA,echo=TRUE}
sum(complete.cases(act)==FALSE)
```
For removing the effect of missing data in the base, I decided to substitute it 
with the respective median value

```{r substitution,echo=TRUE}
sust<-median(act$steps,na.rm = TRUE)
y<-act

for(i in 1:length(y$steps))
if(is.na(y$steps[i])){
  y$steps[i]<-sust
}
```

At the same time I generate a new base with the data replaced called y.
The histogram required and the means and medians are presented in the next block
```{r hist,echo=TRUE}
hist(y$steps,col="blue")
y[,mean(steps),by=date]
y[,median(steps),by=date]
```

The values replaced does not seem to generate a big chance in tendences.

## Are there differences in activity patterns between weekdays and weekends?
Plot required with the differences between the days is presented

```{r week,echo=TRUE}

for(i in 1:length(y$date)){
  if(weekdays(y$date[i])%in%c("sábado","domingo")){
    y$wd[i]<-"weekend"
  }else{
    y$wd[i]<-"weekday"
  }
}

y$wd<-factor(y$wd)

par(mfrow=c(2,1),mar=c(2,2,2,1))
plot(y[wd=="weekday",mean(steps),by=interval]$V1,type="l",col="red",
     main="Weekday")
plot(y[wd=="weekend",mean(steps),by=interval]$V1,type="l",col="blue",
     main="Weekend")
dev.off()
```

